---
title: "Weather Forcaset"
author: "Haoran He"
date: "2022/11/28"
output:
  html_document: default
  word_document: default
---

```{r}
library(lubridate)
library(tseries) 
library(TSA)
library(forecast)
library(ggplot2)
```


```{r}
df <- read.csv("DailyDelhiClimate.csv",header = TRUE)
colnames(df) <- c("Date", "meantemp")
date<-df$Date
newDate<-format(as.Date(date),format="%Y-%m")
Meantemp<-data.frame(df$meantemp)
df2 = data.frame(newDate,Meantemp)
Meantemp <- ts(data=Meantemp, start=c(2013,1),end=c(2017,4),frequency=12)
ts.plot(Meantemp)
```
```{r}
tseries::adf.test(Meantemp)
```

Ho: non-stationary
H1:stationary
If we reject Ho, the model is no more differencing.
```{r}
x=diff(Meantemp)
tseries::adf.test(x)

plot(x)
```
Since P-value bigger than 0.05, .
This time p-value lower than alpha=0.05


if data show AR or MA trend
```{r}
acf(x, 30)
pacf(x, 30)
```
From the graph, sample ACF appears to be more significant.

EACF
```{r}
eacf(x)
```

```{r}
model110 = arima(x, order= c(1,1,0))
model111 = arima(x, order= c(1,1,1))
model011 = arima(x, order= c(0,1,1))
model012 = arima(x, order= c(0,1,2))
model100 = arima(x, order= c(1,0,0))
model213 = arima(x, order= c(2,1,3))

model110$loglik
model111$loglik
model011$loglik 
model012$loglik 
model100$loglik
model213$loglik
```
From the log-likelihoods, the 3 highest values came from models (1,0), (0,2)
and (2,3), but we choose model ARMA(1,0) because our ACF and PACF plots suggests a AR(1) or MA(1) model.


AIC
```{r}
model110 = arima(x, order= c(1,1,0),method="ML")
model111 = arima(x, order= c(1,1,1),method="ML")
model011 = arima(x, order= c(0,1,1),method="ML")
model012 = arima(x, order= c(0,1,2),method="ML")
model100 = arima(x, order= c(1,0,0),method="ML")
model213 = arima(x, order= c(2,1,3),method="ML")

model110$aic
model111$aic
model011$aic
model012$aic
model100$aic
model213$aic
```
Model100 has lowest aic.



**Parameter estimation**
```{r}
model110 = arima(x, order= c(1,0,0))
model110

```
Using the estimated parameter in the table, we can get the function of the model.

```{r}
model100css = arima(x,order= c(1,0,0),method="CSS")
model100ml = arima(x,order= c(1,0,0),method="ML")
model100css
model100ml
```

Residual
```{r}
res=residuals(arima(x,order=c(1,0,0)))
plot(res)
plot(y=res,x=zlag(res,1))
plot(y=res,x=zlag(res,2))
acf(res)
qqnorm(res)
qqline(res)
```
The residual plot seems completely random from ARMA(1,2) model, additionally I checked the lag1 and lag2 correlation plots for the residual, both plots show a horizontal elliptical shape, which means the slope is probabaly 0, indicatiing no correlation between residual and lags of residual. Then I also ploted the ACF of the residuals, to check if there's any time-series pattern left in the residual, all auto-correlations are within the non-rejection range. 


```{r}
train <- subset(Meantemp, end=length(Meantemp)*0.70)
test <- subset(Meantemp, start=length(Meantemp)*0.70)
model <- arima(train, order = c(1,0,0), seasonal = list(order = c(0,1,0),period =12), method = "ML")
forecast_data <-forecast(model,h=24,level=90) %>%
  autoplot() + autolayer(test)
plot(forecast_data,xlab = "Time", ylab = "Meantemp",main = "forecasting_data for Meantemp") 
```

